Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-12-23 manuscript.Rmd
Update discussion

\begin{align}
\text{Inverse frequency weighting of overlap words = }\sum_{i=1}^k
\frac{log(f(t_i, d1) + f(t_i, d2) + 1)}{log(f(t_i) + 1)}
\end{align}

Equation 1 identifies the inverse frequency weighting of each overlapping word ti between two tweets d1 and d2, relative to the word’s frequency in the corpus  D, for all overlapping words t1...k between the two tweets.

\begin{align}
\text{Inverse frequency weighting of all words = }\sum_{i=1}^n \frac{log(f(t_i,
d1) + f(t_i, d2) + 1)}{log(f(t_i) + 1)}
\end{align}

Equation 2 then calculates the inverse frequency of each word ti  in tweets d1 and d2, relative to the word’s frequency in the corpus D, for all words t1...n that appear in either tweet.

@@

To perform CRQA, the time series must be converted to the same scale, while
still preserving a sampling rate that respects the native timescale of the
system [cf. Nyquist frequency in signal processing; @grenander1959probability].
Because the social cohesion and event data have different scales and values, we
converted each time series (e.g., Figure \@ref(fig:raw-ts)) into deciles (e.g.,
Figure \@ref(fig:deciled-ts)) to analyze the relative dynamics of different
levels of social cohesion and counts of all, positive, and negative events on
each day. We implemented CRQA using the `crqa` package [@R-crqa] for R
[@R-base].

@@

Using the real deciles of social cohesion, count of all events, count of
positive events, and count of negative events, we conducted a series of
permutation tests (one for the comparison between social cohesion and each type
of event count) to determine statistical difference from chance for each CRQA
metric. Using the `sample` function in base R [@R-base], $1000$ permuted time
series were created for each of the real time series. CRQA was then run $1000$
times for social cohesion and count of all events, social cohesion and count of
positive events, and social cohesion and count of negative events using the
permuted time series. This resulted in $1000$ results for each of the CRQA
metrics for each combination of time series. Statistical significance from
chance was set at the 95th and 99th percentiles of the metrics from the permuted
CRQ analyses. Given that a permutation test is similar to bootstrapping but does
not include replacement, all points were included in each CRQA analysis, leaving
RR constant across all permutations. Thus, RR is not included in the inferential
tests.

@@

#### Parameters
In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, the *window size* will determine how many observations
are used in each CRQA analysis. We conducted WCRQA with a window size of $14$ to
investigate the dynamics over two weeks and allow for the capture of more
structure and thus a more meaningful interpretation of DET.

@@

### Syria as Source and Target
By adding Syria as a source in the data filtering, the event dataset increased in size
by 1,690 samples. These additional observations are times where Syria initiated
an event with another country. Events initiated by Syria but directed internally
within the country were already captured by the target-only filtering. Taking outwardly directed events into account dramatically changed the coupling between all events and social cohesion. There is no longer a significant number of shared trajectories or line lengths, indicating that the cohesion on Twitter is not tied to the frequency of overall daily events. Since source events are at the root of this difference, it is plausible that Syrians were focused on the target events as they are immediately impacting them, and the addition of the source events now overshadows the relationship between target events and social cohesion online. In contrast, there was a slight increase in coupling strength between positive events and social cohesion that once again can only be due to the externally directed events. These two time series showed greater stability in their coupling (reflected in L). But the significant ENTR metric suggests that there is more variability in the coupling, and thus we do not feel that the source events impacted the positive event coupling notably, but rather broke up the trajectories formed by the target events. Although the coupling between social cohesion with negative events were largely unchanged, there was a slight decrease in coupling strength (reflected in maxL) and increase in complexity of shared trajectories (reflected in rENTR). Similarly to that of positive events, this change may not be reflective of source events themselves, but rather a disruption of the relationship with target events.

@@

### Syria as Target
The time period of differing dynamics noted in the cross-recurrence plots---from
observation 32 through 47---appears to be captured by the above-chance RR values
in WCRQA of social cohesion and count of all events (Figure
\@ref(fig:plot-wcrqa-targ-all)). However, DET for these periods only
intermittently rises above chance; it more consistently reaches statistical
significance only at the end of the window through the end of the observation
period, following a spike in the RR between negative event count and social
cohesion. During this period, the modal ICEWS event for each day (Figure
\@ref(fig:plot-mode-event)) was almost consistently -10, the most violent event
category. This supports our first hypothesis.

@@

Surprisingly---and contrary to our second hypothesis---we saw consistent
coupling between positive event count and social cohesion. Although both had a
downward trajectory throughout the time period under consideration, RR was at or
above chance in nearly all windows, and DET showed much more consistent shared
trajectories than negative events. This may suggest that Syrian Twitter
community was reacting more cohesively to the positive events aimed at Syria
from the international community; this interpretation is supported by the
attenuation of the effect in the analyses that include Syria as both a source
and a target.

### Syria as Source and Target
WCRQA on the source-and-target-filtered data showed similar patterns of RR
between the count of all events and social cohesion (as observed in the
target-only data), though the values increased. This suggests that large numbers
on a given day match increased levels of social cohesion online. However, the
results for DET changed: While the majority of statistically significant DET
windows occurred toward the end of the observation period for target-only data
(creating a positive trendline), the source-and-target data showed a negative
trendline for DET.

The divergences in the all-event count data for these two datasets underscores
the differences between RR and DET. As noted earlier, DET captures
*trajectories*---in this case, similar to the impact of actions and social
cohesion carrying across time. The same number of points (RR) can fall in
different patterns (DET), providing important information about the structure of
the system in question.

WCRQA metrics for positive-only event counts with social cohesion showed similar
patterns in RR and DET as the target-only data, but the precise values of DET
and RR were diminished. This may support the ideas suggested in the CRQA
section, but again, only future work including the *content* of those tweets can
directly speak to that explanation.

Finally, the coupling between negative-only event counts and social cohesion
also remained relatively similar but demonstrated some important differences.
First, we see a greater number of windows at above-chance levels for RR,
including an interesting spike around window 32 that suggests a negative
externally focused action taken by Syria prompted a sharp rise in social
cohesion. (This would not have been captured by the target-only dataset.)
Second---like with the all-event data---we see different structuring of the data
through differences in the patterns of DET; above-chance DET only occurrs within
the middle of the observation period, around a time of increased strife (as
noted above).

@@

This methodology should be of great interest to social scientists investigating
social mobilization. It has potential applications for gender studies and
the #MeToo movement and for the relationship between social protest and repression
during the Black Lives Matter protests. Fundamentally, this methodological
approach can help us better understand the human social mechanisms underlying
social media and its influence on human behavior in real life. Given that many
governments enact repressive measures when they determine that collective action
by citizens is imminent, this methodology should help to improve the predictive
power of using social media data in estimating repressive behavior. If social
media messages are increasing in cohesion, this may represent a threat to civil
and political liberties of citizens engaging in collective action.

>>>>

\begin{align}
\text{Inverse frequency weighting of overlap words = }\sum_{i=1}^k
\frac{log(f(t_i, d_1) + f(t_i, d_2) + 1)}{log(f(t_i) + 1)}
\end{align}

Equation 1 identifies the inverse frequency weighting of each overlapping word ti between two tweets d_1 and d_2, relative to the word’s frequency in the corpus D, for all overlapping words t1...k between the two tweets.

\begin{align}
\text{Inverse frequency weighting of all words = }\sum_{i=1}^n \frac{log(f(t_i,
d_1) + f(t_i, d_2) + 1)}{log(f(t_i) + 1)}
\end{align}

Equation 2 then calculates the inverse frequency of each word ti  in tweets d_1 and d_2, relative to the word’s frequency in the corpus D, for all words t1...n that appear in either tweet.

@@

To perform CRQA, the time series must be converted to the same scale, while
still preserving a sampling rate that respects the native timescale of the
system [cf. Nyquist frequency in signal processing; @grenander1959probability]. In other words, a system that changes every day must be sampled more often than a system that changes every year and less often than a system that changes every hour. However, to compare two time series with CRQA, the two time series must be sampled at the same frequency as one another.
Because the social cohesion and event data have different scales and values, we
converted each time series (e.g., Figure \@ref(fig:raw-ts)) into deciles (e.g.,
Figure \@ref(fig:deciled-ts)) to analyze the relative dynamics of different
levels of social cohesion and counts of all, positive, and negative events on
each day. We implemented CRQA using the `crqa` package [@R-crqa] for R
[@R-base].

@@

Using the real deciled time series of social cohesion, count of all events, count of
positive events, and count of negative events, we conducted a series of
permutation tests (one for the comparison between social cohesion and each type
of event count) to determine statistical difference from chance for each CRQA
metric. Using the `sample` function in base R [@R-base], $1000$ permuted time
series were created for each of the real time series. CRQA was then run $1000$
times for social cohesion and count of all events, social cohesion and count of
positive events, and social cohesion and count of negative events using the
permuted time series. This resulted in $1000$ results for each of the CRQA
metrics for each combination of time series. Statistical significance from
chance was set at the 95th and 99th percentiles of the metrics from the permuted
CRQ analyses. Given that a permutation test is similar to bootstrapping but does
not include replacement, all points were included in each CRQA analysis, leaving
RR constant across all permutations. Thus, RR is not included in the inferential
tests.

@@

#### Parameters
In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, the *window size* will determine how many observations
are used in each CRQA analysis. We conducted WCRQA with a window size of $14$ to
investigate the dynamics over two weeks and allow for the capture of more
structure and thus a more meaningful interpretation of DET. A window size smaller than 14 would limit the possibilities for shared trajectories and make determinism uninterpretable, and window sizes larger than 14 would increasingly overshadow the fine-grain dynamics that we want to investigate.

@@

### Syria as Source and Target
By adding Syria as a source in the data filtering, the event dataset increased in size
by N [fix number here] samples. These additional observations are times where Syria initiated
an event with another country. Events initiated by Syria but directed internally
within the country were already captured by the target-only filtering. Taking outwardly directed events into account dramatically changed the coupling between all events and social cohesion. There is no longer a significant number of shared trajectories or line lengths, indicating that the cohesion on Twitter is not tied to the frequency of overall daily events. Since source events are at the root of this difference, it is plausible that Syrians were focused on the target events as they are immediately impacting them, and the addition of the source events now overshadows the relationship between target events and social cohesion online. In contrast, there was a slight increase in coupling strength between positive events and social cohesion that once again can only be due to the externally directed events. These two time series showed greater stability in their coupling (reflected in L). But the significant ENTR metric suggests that there is more variability in the coupling, and thus we do not feel that the source events impacted the positive event coupling notably, but rather broke up the trajectories formed by the target events. Although the coupling between social cohesion with negative events were largely unchanged, there was a slight decrease in coupling strength (reflected in maxL) and increase in complexity of shared trajectories (reflected in rENTR). Similarly to that of positive events, this change may not be reflective of source events themselves, but rather a disruption of the relationship with target events.

@@

### Syria as Target
For the RR values in WCRQA of social cohesion and count of all events (Figure
\@ref(fig:plot-wcrqa-targ-all)), we see both an early and late above-chance peak. However, DET for these periods only rises above chance for the later peak. During this later period, the modal ICEWS event for each day (Figure
\@ref(fig:plot-mode-event)) was almost consistently -10, the most violent event
category. This supports our first hypothesis, that overall online cohesion may reflect the frequency of events occurring on the ground, particularly those that are violent in nature.

@@

///

///

@@

This methodology should be of great interest to social scientists investigating
social mobilization. It has potential applications for gender studies and
the #MeToo movement and for the relationship between social protest and repression
during the Black Lives Matter protests. This methodological
approach can help us better understand the fundamental human social mechanisms underlying
social media and its influence on human behavior in real life. Given that many
governments enact repressive measures when they determine that collective action
by citizens is imminent, this methodology should help to improve the predictive
power of using social media data in estimating repressive behavior. If social
media messages are increasing in cohesion, this may represent a threat to civil
and political liberties of citizens engaging in collective action.
